---
title: "In Class exercise 4"
author: "Brian Lim"
date: "September 9, 2024"
date-modified: "last-modified"
categories:
  - Analysis
  - R
  - sf
  - tidyverse
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
  cache: true
---

## **4.0 Overview**

A spatio-temporal point process (also called space-time or spatial-temporal point process) is a random collection of points, where each point represents the time and location of an event. Examples of events include incidence of disease, sightings or births of a species, or the occurrences of fires, earthquakes, lightning strikes, tsunamis, or volcanic eruptions.

The analysis of spatio-temporal point patterns is becoming increasingly necessary, given the rapid emergence of geographically and temporally indexed data in a wide range of fields. Several spatio-temporal point patterns analysis methods have been introduced and implemented in R in the last ten years. This chapter shows how various R packages can be combined to run a set of spatio-temporal point pattern analyses in a guided and intuitive way. A real world forest fire events in Kepulauan Bangka Belitung, Indonesia from 1st January 2023 to 31st December 2023 is used to illustrate the methods, procedures and interpretations.

## **4.1 Learning Outcome**

### **4.1.1 The research questions**

The specific question we would like to answer is:

-   are the locations of forest fire in Kepulauan Bangka Belitung spatial and spatio-temporally independent?

-   if the answer is NO, where and when the observed forest fire locations tend to cluster?

### **4.1.2 The data**

For the purpose of this exercise, two data sets are used, they are:

-   forestfires, a csv file provides locations of forest fire detected from the Moderate Resolution Imaging Spectroradiometer (MODIS) sensor data. The data are downloaded from [Fire Information for Resource Management System](https://firms.modaps.eosdis.nasa.gov/download/). For the purpose of this exercise, only forest fires within Kepulauan Bangka Belitung will be used.

-   Kepulauan_Bangka_Belitung, an ESRI shapefile showing the sub-district (i.e. [kelurahan](https://en.wikipedia.org/wiki/Villages_of_Indonesia)) boundary of Kepulauan Bangka Belitung. The data set was downloaded from [Indonesia Geospatial](https://www.indonesia-geospasial.com/2023/05/download-shapefile-batas-administrasi.html) portal. The original data covers the whole Indonesia. For the purpose of this exercise, only sub-districts within Kepulauan Bangka Belitung are extracted.

## **4.2 Installing and Loading the R packages**

For the purpose of this study, five R packages will be used. They are:

-   rgdal for importing geospatial data in GIS file format such as shapefile into R and save them as Spatial\*DataFrame,

-   maptools for converting Spatial\* object into ppp object,

-   raster for handling raster data in R,

-   sparr provides functions to estimate fixed and adaptive kernel-smooth spatial relative risk surfaces via the density-ratio method and perform subsequent inferences,

-   spatstat for performing Spatial Point Patterns Analysis such as kcross, Lcross, etc., and

-   tmap for producing cartographic quality thematic maps.

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

## **4.3 Importing data into R**
### **4.3.1 Importing and Preparing Forest Fire shapeful**
```{r}
kbb_sf <- st_read(dsn = "data/In-class_Ex04", 
                  layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>% # dropping the Z-value
  st_transform(crs = 32748)
```

```{r}
summary(kbb_sf)
```
Furthermore we create an `owin` object from the `sf` data type
```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

To ensure that the output is an `owin` object
```{r}
class(kbb_owin)
```

### **4.3.2 Importing and Preparing Forest Fire data**

Next we will import the forest data set into R:
```{r}
fire_sf <- read_csv("data/In-class_Ex04/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"), 
           crs = 4326) %>%
  st_transform(crs = 32748)
```
Because **ppp** object only accept numerical or character as mark. The code below is used to transform **acq_date** data type to numeric.

```{r}
fire_sf <- fire_sf %>%
  mutate("DayofYear" = yday(acq_date)) %>%
  mutate("Month_num" = month(acq_date)) %>%
  mutate("Month_fac" = month(acq_date, label = TRUE, abbr = FALSE))

summary(fire_sf)
```
## **4.4 Visualising the Fire Points**
```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
  tm_shape(fire_sf) +
  tm_dots()
```
We will then prepare a point symbol map showing the monthly geographic distribution of forest fires in 2023. The map should look similar to the figure below.
```{r}
tm_shape(kbb_sf)+
  tm_polygons()+
  tm_shape(fire_sf)+
  tm_dots(size = 0.1)+
  tm_facets(by="Month_fac", free.coords = FALSE, drop.units = FALSE)
```
## **4.5 Extracting forest fire by month**

The code chunk below is used to remove the unwanted fields from the `fire_sf` sf data frame. This is because as.ppp() only need the mark field and geometry field from the input sf data frame.
```{r}
fire_month <- fire_sf %>%
  select(Month_num)
```

### **4.5.1 Creating ppp**
The code chunk below is used to derive a ppp object called fire_month from fire_month sf data frame
```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```
### **4.5.2 Creating Owin object**
The code chunk below is used to combine origin_am_ppp and am_owin_objects into one.
```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```
Computing Spatio-Temporal KDE
Next, `spattemp.density()` of `sparr` package is used to compute the STKDE
```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```
In the code chunk below, `plot()` of R base is used to get the KDE for between July 2023 - December 2023
```{r}
tims <- c(7, 8, 9, 10, 11, 12)
par(mfcol=c(2, 3))
for(i in tims){
  plot(st_kde, i, 
       override.par=FALSE,
       fix.range=TRUE,
       main=paste("KDE at month", i))
}
```
```{r}
kde_yday <- spattemp.density(
  fire_yday_owin)

summary(kde_yday)
```

















