---
title: "In Class exercise 11"
author: "Brian Lim"
date: "October 21, 2024"
date-modified: "last-modified"
categories:
  - Analysis
  - R
  - sf
  - tidyverse
  - tmap
  - httr
  - performance
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true  
  cache: true
---

## Loading the R packages

```{r}
pacman::p_load(tidyverse, sf, tmap, httr, performance)
```


The code chunk below imports multiple csv files in a specified folder and append them into a single tibble data frame.
```{r}
folder_path <- "data/In-class_Ex12/aspatial"
file_list <- list.files(path = folder_path, 
                        pattern = "^realis.*\\.csv$", 
                        full.names = TRUE)

realis_data <- file_list %>%
  map_dfr(read_csv)
```
The following code chunk converts values in Sale Date field from character to numerical date format, and - extracting resale and condominium transaction records.
```{r}
condo_resale <- realis_data %>%
  mutate(`Sale Date` = dmy(`Sale Date`)) %>%
  filter(`Type of Sale` == "Resale" &
           `Property Type` == "Condominium")
```

## Data preparation
```{r}
postcode <- unique(condo_resale$`Postal Code`)
```

```{r}
url <- "https://onemap.gov.sg/api/common/elastic/search"
found <- data.frame()
not_found <- data.frame()

for (postcode in postcode){
  query <- list('searchVal'=postcode, 'returnGeom'='Y', 
                'getAddrDetails'='Y', 'pageNum'='1')
  res <- GET(url, query=query)
  if ((content(res)$found)!=0){
    found <- rbind(found, data.frame(content(res))[4:13])
  } else {not_found = data.frame(postcode)
  }
}
```

## Tidying up field names
```{r}
found <- found %>%
  select(c(6:8)) %>%
  rename(POSTAL = `results.POSTAL`,
         XCOORD = `results.X`,
         YCOORD = `results.Y`)
```

The following chunk of code is to join `condo_resale` and `found`
```{r}
condo_resale_geocoded = left_join(
  condo_resale, found, 
  by = c('Postal Code' = 'POSTAL'))
```

The following chunk of code is to convert `condo_resale_geocoded` from **tibble** data frame to **sf** point feature data frame
```{r}
condo_resale_sf <- st_as_sf(condo_resale_geocoded, 
                            coords = c("XCOORD",
                                       "YCOORD"),
                            crs=3414)
```

### Cleaning spatial data
The code chunk below is used to check if there are overlapping point features.
```{r}
overlapping_points <- condo_resale_sf %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```

In the code code chunk below, `st_jitter()` of sf package is used to move the point features by 5m to avoid overlapping point features.
```{r}
condo_resale_sf <- condo_resale_sf %>%
  st_jitter(amount = 2)
```